---
title: "visual_emotion_processing_psychopathology"
output: html_document
date: "2025-11-25"
---

# Main steps
 - part 1 - general setup
 - part 2 - dissimilarity and total problem score
 - part 3 - dissimilarity and performance
 - part 4 - effect sizes
 - part 5 - performance and total problem score
 - part 6 - demographic table
 - part 7 - mediation
 
note: Pearson correlations calculated by python RSAtoolbox package

author: Yen-Chu Lin

# part 1 - general setup
- (1) load packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lme4)
library(lmerTest)
library(sjPlot)
require(MuMIn)
library(effsize) # for cohen.d
library(tableone)
library(vtable)
library(summarytools)
```

- (2) functions
```{r}
select_condition_and_combine_dataframes_v1 <- function(
    r_visual, r_limbic, r_dorsal, r_ventral, r_frontal, r_default, r_somatic, r_v1, col_name){
  # the function subsets dataframe based on condition selected & combine all dataframes including v1
  # also see select_conditions & combine_networks
  
  # subset data based on col_name
  rvis <- r_visual[c("src_subject_id",col_name)]
  rlim <- r_limbic[c("src_subject_id",col_name)]
  rdor <- r_dorsal[c("src_subject_id",col_name)]
  rven <- r_ventral[c("src_subject_id",col_name)]
  rfro <- r_frontal[c("src_subject_id",col_name)]
  rdef <- r_default[c("src_subject_id",col_name)]
  rsom <- r_somatic[c("src_subject_id",col_name)]
  rv1  <- r_v1[c("src_subject_id",col_name)]
  
  colnames(rvis)[2] <- "vis"
  colnames(rlim)[2] <- "lim"
  colnames(rdor)[2] <- "dor"
  colnames(rven)[2] <- "ven"
  colnames(rfro)[2] <- "fro"
  colnames(rdef)[2] <- "def"
  colnames(rsom)[2] <- "som"
  colnames(rv1)[2]  <- "v1"

  # combine all 7 networks & v1 into one dataframe
  data <- merge(rvis, rlim, by="src_subject_id")
  data <- merge(data, rdor, by="src_subject_id")
  data <- merge(data, rven, by="src_subject_id")
  data <- merge(data, rfro, by="src_subject_id")
  data <- merge(data, rdef, by="src_subject_id")
  data <- merge(data, rsom, by="src_subject_id")
  data <- merge(data, rv1,  by="src_subject_id")
  
  # return data
  return(data)
}

func_merge_df <- function(data_r, fam_site_id, nback) {
  # merge with family ad site ids
  data_full <- merge(data_r, fam_site_id, by="src_subject_id")
  # merge with nback
  data_full <- merge(data_full, nback, by="src_subject_id")
  
  return(data_full)
}

vector_to_factor <- function(data_long) {
  data_long$src_subject_id <- as.factor(data_long$src_subject_id)
  data_long$condition      <- as.factor(data_long$condition)
  data_long$rel_family_id  <- as.factor(data_long$rel_family_id)
  data_long$site_id_l      <- as.factor(data_long$site_id_l)
  data_long$network        <- as.factor(data_long$network)
  return(data_long)
}
```

- (3) inputs
```{r}
# file path
r_file_path    <- "/Volumes/FABlab drive 2/Processed data/RSA_Pearson_correlations/"
abcd_file_path <- "/Volumes/FABlab drive 2/Data/abcd-data-release-4.0/"
subj_list_path <- "/Volumes/FABlab drive 2/Processed data/"

# dissimilarity (Pearson correlation files)
file1  <- "visual_pearson_corr.csv"
file2  <- "limbic_pearson_corr_0back.csv"
file3  <- "limbic_pearson_corr_2back.csv"
file4  <- "doratt_pearson_corr_0back.csv"
file5  <- "doratt_pearson_corr_2back.csv"
file6  <- "venatt_pearson_corr_0back.csv"
file7  <- "venatt_pearson_corr_2back.csv"
file8  <- "sommot_pearson_corr_0back.csv"
file9  <- "sommot_pearson_corr_2back.csv"
file10 <- "control_pearson_corr_0back.csv"
file11 <- "control_pearson_corr_2back.csv"
file12 <- "default_pearson_corr_0back.csv"
file13 <- "default_pearson_corr_2back.csv"
file14 <- "v1_pearson_corr.csv"

# CBCL scores
cbcl_filename <- "abcd_cbcls01.txt"
cbcl_col_select <- 
    c("subjectkey",
    "cbcl_scr_dsm5_adhd_t","cbcl_scr_dsm5_adhd_r",
    "cbcl_scr_syn_attention_t","cbcl_scr_syn_attention_r",
    "cbcl_scr_syn_external_t","cbcl_scr_syn_external_r",
    "cbcl_scr_syn_internal_t","cbcl_scr_syn_internal_r",
    "cbcl_scr_syn_totprob_t","cbcl_scr_syn_totprob_r"
    )

# N-back behavioral data (accuracy)
naback_behav_filename <- "abcd_mrinback02.txt"
nback_col_select <- 
  c("subjectkey",
    "tfmri_nb_all_beh_c0bpf_rate","tfmri_nb_all_beh_c0bnf_rate","tfmri_nb_all_beh_c0bngf_rate",
    "tfmri_nb_all_beh_c2bpf_rate","tfmri_nb_all_beh_c2bnf_rate","tfmri_nb_all_beh_c2bngf_rate",
    "tfmri_nb_all_beh_c0b_rate","tfmri_nb_all_beh_c2b_rate"
    )

# list of clean subjects
subj_filename  <- "cleanedsubjectkeys.csv"

# family income, caregiver education, and child race/ethnicity
pdemo_filename <- "pdem02.txt"
pdemo_col_select <- 
    c("subjectkey","demo_prnt_ed_v2","demo_prtnr_ed_v2",
      "demo_prnt_income_v2", "demo_prtnr_income_v2","demo_comb_income_v2",
      "demo_race_a_p___10", "demo_race_a_p___11", "demo_race_a_p___12", "demo_race_a_p___13",
      "demo_race_a_p___14", "demo_race_a_p___15", "demo_race_a_p___16", "demo_race_a_p___17", 
      "demo_race_a_p___18", "demo_race_a_p___19", "demo_race_a_p___20", "demo_race_a_p___21", 
      "demo_race_a_p___22", "demo_race_a_p___23", "demo_race_a_p___24", "demo_race_a_p___25", 
      "demo_race_a_p___77", "demo_race_a_p___99","demo_ethn_v2")
```

- (4) load files
```{r}
# load the dissimilarity files
r_vis    <- read.csv(paste(r_file_path, file1,  sep=""), header = TRUE)
r_lim_0b <- read.csv(paste(r_file_path, file2,  sep=""), header = TRUE)
r_lim_2b <- read.csv(paste(r_file_path, file3,  sep=""), header = TRUE)
r_dor_0b <- read.csv(paste(r_file_path, file4,  sep=""), header = TRUE)
r_dor_2b <- read.csv(paste(r_file_path, file5,  sep=""), header = TRUE)
r_ven_0b <- read.csv(paste(r_file_path, file6,  sep=""), header = TRUE)
r_ven_2b <- read.csv(paste(r_file_path, file7,  sep=""), header = TRUE)
r_som_0b <- read.csv(paste(r_file_path, file8,  sep=""), header = TRUE)
r_som_2b <- read.csv(paste(r_file_path, file9,  sep=""), header = TRUE)
r_fro_0b <- read.csv(paste(r_file_path, file10, sep=""), header = TRUE)
r_fro_2b <- read.csv(paste(r_file_path, file11, sep=""), header = TRUE)
r_def_0b <- read.csv(paste(r_file_path, file12, sep=""), header = TRUE)
r_def_2b <- read.csv(paste(r_file_path, file13, sep=""), header = TRUE)
r_v1     <- read.csv(paste(r_file_path, file14, sep=""), header = TRUE)
r_vis_0b <- r_vis[c("src_subject_id","X0happy0fear","X0happy0neu","X0neu0fear")]
r_vis_2b <- r_vis[c("src_subject_id","X2happy2fear","X2happy2neu","X2neu2fear")]
r_v1_0b  <- r_v1[c("src_subject_id","X0happy0fear","X0happy0neu","X0neu0fear")]
r_v1_2b  <- r_v1[c("src_subject_id","X2happy2fear","X2happy2neu","X2neu2fear")]

# load cbcl
abcd_cbcl <- read.delim(paste(abcd_file_path, cbcl_filename, sep=""), header = TRUE) %>% filter(eventname == "2_year_follow_up_y_arm_1")
abcd_cbcl <- abcd_cbcl[cbcl_col_select]
abcd_cbcl$subjectkey <- gsub("NDAR_", "", abcd_cbcl$subjectkey)
names(abcd_cbcl)[names(abcd_cbcl) == 'subjectkey'] <- 'src_subject_id'

# load the file with family_id and subset the columns we need 
acspsw  <- read.delim(paste(abcd_file_path, "acspsw03.txt", sep=""), header = TRUE)  %>% 
  slice(-1) %>%
  filter(eventname == "baseline_year_1_arm_1") %>% # only has data from baseline & year 1 
  subset(select = c("subjectkey", "rel_family_id"))

# load the file with site_id and subset the columns we need
abcd_site <- read.delim(paste(abcd_file_path, "abcd_lt01.txt", sep=""), header = TRUE) %>% 
  slice(-1) %>%
  filter(eventname == "2_year_follow_up_y_arm_1") %>%
  subset(select = c("subjectkey", "site_id_l","interview_age","sex")) 

# combine family id and site id
fam_site_id <- merge(acspsw, abcd_site, by = "subjectkey")
fam_site_id$subjectkey <- gsub("NDAR_", "", fam_site_id$subjectkey)
names(fam_site_id)[names(fam_site_id) == 'subjectkey'] <- 'src_subject_id'

# load a clean list of subject to use in the neuroimaging analysis
subj_list <- read.csv(paste(subj_list_path, subj_filename, sep=""), header = FALSE) %>% 
  slice(-1)

# load N-back behavioral data
nback <- read.delim(paste(abcd_file_path, naback_behav_filename, sep=""), header = TRUE) %>% filter(eventname == "2_year_follow_up_y_arm_1")
nback <- nback[nback_col_select] %>% na.omit()
nback$subjectkey <- gsub("NDAR_", "", nback$subjectkey)
names(nback)[names(nback) == 'subjectkey'] <- 'src_subject_id'

# load family income and caregiver education data
abcd_pdem <- read.delim(paste(abcd_file_path, pdemo_filename, sep=""), header = TRUE) %>% 
  slice(-1) %>%
  select(all_of(pdemo_col_select))
abcd_pdem <- abcd_pdem[pdemo_col_select] %>% na.omit()
abcd_pdem$subjectkey <- gsub("NDAR_", "", abcd_pdem$subjectkey)
names(abcd_pdem)[names(abcd_pdem) == 'subjectkey'] <- 'src_subject_id'
```

- (5) preprocess dataframes
```{r}
# dissimilarity
## SUBSETTING DATA BASED ON CONDITION SELECTED
### 0-back ###
col_name1 <- "X0happy0fear"
col_name2 <- "X0happy0neu"
col_name3 <- "X0neu0fear"

# subset and combine dataframes
data_hf_0b <- select_condition_and_combine_dataframes_v1(
  r_vis_0b, r_lim_0b, r_dor_0b, r_ven_0b, r_fro_0b, r_def_0b, r_som_0b, r_v1_0b, col_name1)
data_hn_0b <- select_condition_and_combine_dataframes_v1(
  r_vis_0b, r_lim_0b, r_dor_0b, r_ven_0b, r_fro_0b, r_def_0b, r_som_0b, r_v1_0b, col_name2)
data_fn_0b <- select_condition_and_combine_dataframes_v1(
  r_vis_0b, r_lim_0b, r_dor_0b, r_ven_0b, r_fro_0b, r_def_0b, r_som_0b, r_v1_0b, col_name3)

### 2-back ###
col_name4 <- "X2happy2fear"
col_name5 <- "X2happy2neu"
col_name6 <- "X2neu2fear"

# subset and combine dataframes
data_hf_2b <- select_condition_and_combine_dataframes_v1(
  r_vis_2b, r_lim_2b, r_dor_2b, r_ven_2b, r_fro_2b, r_def_2b, r_som_2b, r_v1_2b, col_name4)
data_hn_2b <- select_condition_and_combine_dataframes_v1(
  r_vis_2b, r_lim_2b, r_dor_2b, r_ven_2b, r_fro_2b, r_def_2b, r_som_2b, r_v1_2b, col_name5)
data_fn_2b <- select_condition_and_combine_dataframes_v1(
  r_vis_2b, r_lim_2b, r_dor_2b, r_ven_2b, r_fro_2b, r_def_2b, r_som_2b, r_v1_2b, col_name6)

# sanity check
# exclude subjects with NA for dissimilarity
# dim(data_hf_0b) # 4952    9
# dim(data_hf_2b) # 4952    9
# dim(data_hn_0b) # 4952    9
# dim(data_hn_2b) # 4952    9
# dim(data_fn_0b) # 4952    9
# dim(data_fn_2b) # 4952    9

# check completeness 
# dim(data_hf_0b[complete.cases(data_hf_0b), ]) # 4952    9
# dim(data_hf_2b[complete.cases(data_hf_2b), ]) # 4952    9
# dim(data_hn_0b[complete.cases(data_hn_0b), ]) # 4952    9
# dim(data_hn_2b[complete.cases(data_hn_2b), ]) # 4952    9
# dim(data_fn_0b[complete.cases(data_fn_0b), ]) # 4952    9
# dim(data_fn_2b[complete.cases(data_fn_2b), ]) # 4952    9

# head(data_hf_0b)
```

# part 2 - dissimilarity and CBCL score 
- (1) process datasets for dissimilarity and total problems score
```{r}
# merge dataframes
### 0-back ###
# (1) add a column to specify the condition
data_hf_0b$condition <- "hf" 
data_hn_0b$condition <- "hn"
data_fn_0b$condition <- "fn"

# (2) merge distance with family id and site id
data_hf0b <- func_merge_df(data_hf_0b, fam_site_id, nback)
data_hn0b <- func_merge_df(data_hn_0b, fam_site_id, nback)
data_fn0b <- func_merge_df(data_fn_0b, fam_site_id, nback)

# (3) merge with cbcl
data_hf0b <- merge(data_hf0b, abcd_cbcl, by="src_subject_id")
data_hn0b <- merge(data_hn0b, abcd_cbcl, by="src_subject_id")
data_fn0b <- merge(data_fn0b, abcd_cbcl, by="src_subject_id")

# (4) convert the data into long format
colname_list <- c("vis","lim","dor","ven","fro","def","som","v1")
data_hf0b_long <- 
  pivot_longer(data_hf0b, cols = all_of(colname_list), names_to = "network", values_to = "dis")
data_hn0b_long <- 
  pivot_longer(data_hn0b, cols = all_of(colname_list), names_to = "network", values_to = "dis")
data_fn0b_long <- 
  pivot_longer(data_fn0b, cols = all_of(colname_list), names_to = "network", values_to = "dis")

# (5) COMBINE ALL THREE CONDITIONS 
data_0back <- rbind(data_hf0b_long, data_hn0b_long, data_fn0b_long)

# (6) make age and CBCL scores numeric
data_hf0b_long <- data_hf0b_long %>%
  mutate_at(c("interview_age","cbcl_scr_syn_totprob_t","cbcl_scr_syn_totprob_r","cbcl_scr_syn_external_r","cbcl_scr_syn_external_t","cbcl_scr_syn_internal_r","cbcl_scr_syn_internal_t"), as.numeric) %>%
  mutate_at(nback_col_select[-1], as.numeric) 
data_hn0b_long <- data_hn0b_long %>%
  mutate_at(c("interview_age","cbcl_scr_syn_totprob_t","cbcl_scr_syn_totprob_r","cbcl_scr_syn_external_r","cbcl_scr_syn_external_t","cbcl_scr_syn_internal_r","cbcl_scr_syn_internal_t"), as.numeric) %>%
  mutate_at(nback_col_select[-1], as.numeric) 
data_fn0b_long <- data_fn0b_long %>%
  mutate_at(c("interview_age","cbcl_scr_syn_totprob_t","cbcl_scr_syn_totprob_r","cbcl_scr_syn_external_r","cbcl_scr_syn_external_t","cbcl_scr_syn_internal_r","cbcl_scr_syn_internal_t"), as.numeric) %>%
  mutate_at(nback_col_select[-1], as.numeric) 
data_0back <- data_0back %>%
  mutate_at(c("interview_age","cbcl_scr_syn_totprob_t","cbcl_scr_syn_totprob_r","cbcl_scr_syn_external_r","cbcl_scr_syn_external_t","cbcl_scr_syn_internal_r","cbcl_scr_syn_internal_t"), as.numeric) %>%
  mutate_at(nback_col_select[-1], as.numeric)

### 2-back ###
# (1) add a column to specify the condition
data_hf_2b$condition <- "hf" 
data_hn_2b$condition <- "hn"
data_fn_2b$condition <- "fn"

# (2) merge distance with family id and site id
data_hf2b <- func_merge_df(data_hf_2b, fam_site_id, nback)
data_hn2b <- func_merge_df(data_hn_2b, fam_site_id, nback)
data_fn2b <- func_merge_df(data_fn_2b, fam_site_id, nback)

# (3) merge with cbcl
data_hf2b <- merge(data_hf2b, abcd_cbcl, by="src_subject_id")
data_hn2b <- merge(data_hn2b, abcd_cbcl, by="src_subject_id")
data_fn2b <- merge(data_fn2b, abcd_cbcl, by="src_subject_id")

# (4) convert the data into long format
colname_list <- c("vis","lim","dor","ven","fro","def","som","v1")
data_hf2b_long <- 
  pivot_longer(data_hf2b, cols = all_of(colname_list), names_to = "network", values_to = "dis")
data_hn2b_long <- 
  pivot_longer(data_hn2b, cols = all_of(colname_list), names_to = "network", values_to = "dis")
data_fn2b_long <- 
  pivot_longer(data_fn2b, cols = all_of(colname_list), names_to = "network", values_to = "dis")

# (5) COMBINE ALL THREE CONDITIONS 
data_2back <- rbind(data_hf2b_long, data_hn2b_long, data_fn2b_long)

# (6) make age and CBCL scores numeric
data_hf2b_long <- data_hf2b_long %>%
  mutate_at(c("interview_age","cbcl_scr_syn_totprob_t","cbcl_scr_syn_totprob_r","cbcl_scr_syn_external_r","cbcl_scr_syn_external_t","cbcl_scr_syn_internal_r","cbcl_scr_syn_internal_t"), as.numeric) %>%
  mutate_at(nback_col_select[-1], as.numeric) 
data_hn2b_long <- data_hn2b_long %>%
  mutate_at(c("interview_age","cbcl_scr_syn_totprob_t","cbcl_scr_syn_totprob_r","cbcl_scr_syn_external_r","cbcl_scr_syn_external_t","cbcl_scr_syn_internal_r","cbcl_scr_syn_internal_t"), as.numeric) %>%
  mutate_at(nback_col_select[-1], as.numeric) 
data_fn2b_long <- data_fn2b_long %>%
  mutate_at(c("interview_age","cbcl_scr_syn_totprob_t","cbcl_scr_syn_totprob_r","cbcl_scr_syn_external_r","cbcl_scr_syn_external_t","cbcl_scr_syn_internal_r","cbcl_scr_syn_internal_t"), as.numeric) %>%
  mutate_at(nback_col_select[-1], as.numeric) 
data_2back <- data_2back %>%
  mutate_at(c("interview_age","cbcl_scr_syn_totprob_t","cbcl_scr_syn_totprob_r","cbcl_scr_syn_external_r","cbcl_scr_syn_external_t","cbcl_scr_syn_internal_r","cbcl_scr_syn_internal_t"), as.numeric) %>%
  mutate_at(nback_col_select[-1], as.numeric)
```

- (2) mixed model - dis vs. cbcl
one model per network
```{r}
vector_to_factor <- function(data_long) {
  data_long$src_subject_id <- as.factor(data_long$src_subject_id)
  data_long$rel_family_id  <- as.factor(data_long$rel_family_id)
  data_long$site_id_l      <- as.factor(data_long$site_id_l)
  return(data_long)
}
data_0back <- vector_to_factor(data_0back)
data_2back <- vector_to_factor(data_2back)

### total problem score ### 
## 0-back ##
# vis
lme2_0back_vis_tol <- lmer(cbcl_scr_syn_totprob_r ~ dis * condition + sex + interview_age + (1|site_id_l/rel_family_id), data = data_0back[data_0back$network == "vis",])
summary(lme2_0back_vis_tol) # 0.940 ; no interactions
# lim
lme2_0back_lim_tol <- lmer(cbcl_scr_syn_totprob_r ~ dis * condition + sex + interview_age + (1|site_id_l/rel_family_id), data = data_0back[data_0back$network == "lim",])
summary(lme2_0back_lim_tol) # 0.487 ; no interactions
# v1
lme2_0back_v1_tol <- lmer(cbcl_scr_syn_totprob_r ~ dis * condition + sex + interview_age + (1|site_id_l/rel_family_id), data = data_0back[data_0back$network == "v1",])
summary(lme2_0back_v1_tol) # 0.0359 * ; no interactions

## 2-back ##
# vis
lme2_2back_vis_tol <- lmer(cbcl_scr_syn_totprob_r ~ dis * condition + sex + interview_age + (1|site_id_l/rel_family_id), data = data_2back[data_2back$network == "vis",])
summary(lme2_2back_vis_tol) # 0.00016 *** ; no interactions
# lim
lme2_2back_lim_tol <- lmer(cbcl_scr_syn_totprob_r ~ dis * condition + sex + interview_age + (1|site_id_l/rel_family_id), data = data_2back[data_2back$network == "lim",])
summary(lme2_2back_lim_tol) # 0.477 ; no interactions
# v1
lme2_2back_v1_tol <- lmer(cbcl_scr_syn_totprob_r ~ dis * condition + sex + interview_age + (1|site_id_l/rel_family_id), data = data_2back[data_2back$network == "v1",])
summary(lme2_2back_v1_tol) # 6.73e-05 *** ; no interactions

### internalization score ### 
## 0-back ##
# vis
lme2_0back_vis_int <- lmer(cbcl_scr_syn_internal_r ~ dis * condition + sex + interview_age + (1|site_id_l/rel_family_id), data = data_0back[data_0back$network == "vis",])
summary(lme2_0back_vis_int) # 0.0145 * ; no interactions
# lim
lme2_0back_lim_int <- lmer(cbcl_scr_syn_internal_r ~ dis * condition + sex + interview_age + (1|site_id_l/rel_family_id), data = data_0back[data_0back$network == "lim",])
summary(lme2_0back_lim_int) # 0.834 ; no interactions
# v1
lme2_0back_v1_int <- lmer(cbcl_scr_syn_internal_r ~ dis * condition + sex + interview_age + (1|site_id_l/rel_family_id), data = data_0back[data_0back$network == "v1",])
summary(lme2_0back_v1_int) # 0.473 ; no interactions

## 2-back ##
# vis
lme2_2back_vis_int <- lmer(cbcl_scr_syn_internal_r ~ dis * condition + sex + interview_age + (1|site_id_l/rel_family_id), data = data_2back[data_2back$network == "vis",])
summary(lme2_2back_vis_int) # 0.325 ; no interactions
# lim
lme2_2back_lim_int <- lmer(cbcl_scr_syn_internal_r ~ dis * condition + sex + interview_age + (1|site_id_l/rel_family_id), data = data_2back[data_2back$network == "lim",])
summary(lme2_2back_lim_int) # 0.970 ; no interactions
# v1
lme2_2back_v1_int <- lmer(cbcl_scr_syn_internal_r ~ dis * condition + sex + interview_age + (1|site_id_l/rel_family_id), data = data_2back[data_2back$network == "v1",])
summary(lme2_2back_v1_int) # 0.624 ; no interactions

### externalization score ### 
## 0-back ##
# vis
lme2_0back_vis_ext <- lmer(cbcl_scr_syn_external_r ~ dis * condition + sex + interview_age + (1|site_id_l/rel_family_id), data = data_0back[data_0back$network == "vis",])
summary(lme2_0back_vis_ext) # 0.168 ; no interactions
# lim
lme2_0back_lim_ext <- lmer(cbcl_scr_syn_external_r ~ dis * condition + sex + interview_age + (1|site_id_l/rel_family_id), data = data_0back[data_0back$network == "lim",])
summary(lme2_0back_lim_ext) # 0.265 ; no interactions
# v1
lme2_0back_v1_ext <- lmer(cbcl_scr_syn_external_r ~ dis * condition + sex + interview_age + (1|site_id_l/rel_family_id), data = data_0back[data_0back$network == "v1",])
summary(lme2_0back_v1_ext) # 0.0217 * ; no interactions

## 2-back ##
# vis
lme2_2back_vis_ext <- lmer(cbcl_scr_syn_external_r ~ dis * condition + sex + interview_age + (1|site_id_l/rel_family_id), data = data_2back[data_2back$network == "vis",])
summary(lme2_2back_vis_ext) # 0.001170 ** ; no interactions
# lim
lme2_2back_lim_ext <- lmer(cbcl_scr_syn_external_r ~ dis * condition + sex + interview_age + (1|site_id_l/rel_family_id), data = data_2back[data_2back$network == "lim",])
summary(lme2_2back_lim_ext) # 0.554 ; no interactions
# v1
lme2_2back_v1_ext <- lmer(cbcl_scr_syn_external_r ~ dis * condition + sex + interview_age + (1|site_id_l/rel_family_id), data = data_2back[data_2back$network == "v1",])
summary(lme2_2back_v1_ext) # 1.01e-05 *** ; no interactions
```

- (3) plot - dissimilarity vs. cbcl
(3.1) cbcl score distributions
```{r}
# preprocess dataset
cbcl_col_list <- 
    c(
      "cbcl_scr_syn_attention_t","cbcl_scr_syn_attention_r",
      "cbcl_scr_syn_external_t","cbcl_scr_syn_external_r",
      "cbcl_scr_syn_internal_t","cbcl_scr_syn_internal_r",
      "cbcl_scr_syn_totprob_t","cbcl_scr_syn_totprob_r"
    )
data_cbcl <- data_hf0b %>%
  mutate_at(cbcl_col_list, as.numeric)

f1_totptob_r <- ggplot(data_cbcl, aes(x=cbcl_scr_syn_totprob_r)) + 
  geom_histogram(binwidth=2, color = "black", fill = "white") +
  scale_y_continuous(expand = c(0,0), limits = c(0,600)) +
  labs(x = "total problems raw score",
       y = "number of subjects") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"), text = element_text(size = 18), aspect.ratio = 1)

f1_ext_r <- ggplot(data_cbcl, aes(x=cbcl_scr_syn_external_r)) + 
  geom_histogram(binwidth=1, color = "black", fill = "white") +
  scale_y_continuous(expand = c(0,0), limits = c(0,1700)) +
  labs(x = "externalizing problems raw score",
       y = "number of subjects") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"), text = element_text(size = 18), aspect.ratio = 1)

f1_int_r <- ggplot(data_cbcl, aes(x=cbcl_scr_syn_internal_r)) + 
  geom_histogram(binwidth=1, color = "black", fill = "white") +
  scale_y_continuous(expand = c(0,0), limits = c(0,1000)) +
  labs(x = "internalizing problems raw score",
       y = "number of subjects") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"), text = element_text(size = 18), aspect.ratio = 1)

f1_totptob_t <- ggplot(data_cbcl, aes(x=cbcl_scr_syn_totprob_t)) + 
  geom_histogram(binwidth=1, color = "black", fill = "white") +
  scale_x_continuous(expand = c(0,0), limits = c(20,95)) +
  scale_y_continuous(expand = c(0,0), limits = c(0,350)) +
  labs(x = "total problems t-score",
       y = "number of subjects") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"), text = element_text(size = 18), aspect.ratio = 1)

f1_ext_t <- ggplot(data_cbcl, aes(x=cbcl_scr_syn_external_t)) + 
  geom_histogram(binwidth=1, color = "black", fill = "white") +
  scale_x_continuous(expand = c(0,0), limits = c(20,95)) +
  scale_y_continuous(expand = c(0,0), limits = c(0,1300)) +
  labs(x = "externalizing problems t-score",
       y = "number of subjects") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"), text = element_text(size = 18), aspect.ratio = 1)

f1_int_t <- ggplot(data_cbcl, aes(x=cbcl_scr_syn_internal_t)) + 
  geom_histogram(binwidth=1, color = "black", fill = "white") +
  scale_x_continuous(expand = c(0,0),limits = c(20,95)) +
  scale_y_continuous(expand = c(0,0), limits = c(0,600)) +
  labs(x = "internalizing problems t-score",
       y = "number of subjects") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"), text = element_text(size = 18), aspect.ratio = 1)

f1_totptob_r
f1_ext_r
f1_int_r
f1_totptob_t
f1_ext_t
f1_int_t

# sanity check
# min(data_0back["cbcl_scr_syn_totprob_r"]) # 0
# max(data_0back["cbcl_scr_syn_totprob_r"]) # 161
# min(data_0back["cbcl_scr_syn_internal_r"]) # 0
# max(data_0back["cbcl_scr_syn_internal_r"]) # 50
# min(data_0back["cbcl_scr_syn_external_r"]) # 0
# max(data_0back["cbcl_scr_syn_external_r"]) # 50
# min(data_2back["cbcl_scr_syn_totprob_r"]) # 0
# max(data_2back["cbcl_scr_syn_totprob_r"]) # 161
# min(data_2back["cbcl_scr_syn_internal_r"]) # 0
# max(data_2back["cbcl_scr_syn_internal_r"]) # 50
# min(data_2back["cbcl_scr_syn_external_r"]) # 0
# max(data_2back["cbcl_scr_syn_external_r"]) # 50

# min(data_2back["cbcl_scr_syn_totprob_t"]) # 24
# max(data_2back["cbcl_scr_syn_totprob_t"]) # 88
# min(data_2back["cbcl_scr_syn_internal_t"]) # 33
# max(data_2back["cbcl_scr_syn_internal_t"]) # 90
# min(data_2back["cbcl_scr_syn_external_t"]) # 33
# max(data_2back["cbcl_scr_syn_external_t"]) # 83

# min(data_0back["dis"]) # 0.00586022
# max(data_0back["dis"]) # 1.757153
# min(data_2back["dis"]) # 0.006938755
# max(data_2back["dis"]) # 1.900827

# head(data_0back)
```

(3.2) subjects with both ext and int problems (check overlaps)
```{r}
# head(data_cbcl)
sum((data_cbcl$cbcl_scr_syn_external_t>=60) & (data_cbcl$cbcl_scr_syn_internal_t>=60)) # 204
length(data_cbcl$src_subject_id[(data_cbcl$cbcl_scr_syn_external_t>=60) & (data_cbcl$cbcl_scr_syn_internal_t>=60)]) # 204
# data_cbcl$src_subject_id[(data_cbcl$cbcl_scr_syn_external_t>=60) & (data_cbcl$cbcl_scr_syn_internal_t>=60)]
```

(3.3) total problems score plot
```{r}
# cbcl_scr_syn_totprob_r
### 0-back ###
# vis
f2_0back_vis_tol <- 
  ggplot(data_0back[data_0back$network == "vis",], aes(cbcl_scr_syn_totprob_r, dis)) + 
  geom_hex() + stat_smooth(color="red", fill="red", method=lm, size=0.5, level=0.999) +
  labs(x="total problems score", y="dissimilarities", title="Vis", fill="count") +
  scale_x_continuous(expand = c(0,0), limits = c(-10,180)) +
  scale_y_continuous(expand = c(0,0), limits = c(-0.1,2.1)) +
  theme_classic() +
  theme(text = element_text(size = 20), 
        axis.text.x = element_text(size = 18), 
        axis.text.y = element_text(size = 18))
# lim
f2_0back_lim_tol <- 
  ggplot(data_0back[data_0back$network == "lim",], aes(cbcl_scr_syn_totprob_r, dis)) + 
  geom_hex() + stat_smooth(color="red", fill="red", method=lm, size=0.5, level=0.999) +
  labs(x="total problems score", y="dissimilarities", title="Lim", fill="count") +
  scale_x_continuous(expand = c(0,0), limits = c(-10,180)) +
  scale_y_continuous(expand = c(0,0), limits = c(-0.1,2.1)) +
  theme_classic() +
  theme(text = element_text(size = 20), 
        axis.text.x = element_text(size = 18), 
        axis.text.y = element_text(size = 18))
# v1
f2_0back_v1_tol <- 
  ggplot(data_0back[data_0back$network == "v1",], aes(cbcl_scr_syn_totprob_r, dis)) + 
  geom_hex() + stat_smooth(color="red", fill="red", method=lm, size=0.5, level=0.999) +
  labs(x="total problems score", y="dissimilarities", title="V1", fill="count") +
  scale_x_continuous(expand = c(0,0), limits = c(-10,180)) +
  scale_y_continuous(expand = c(0,0), limits = c(-0.1,2.1)) +
  theme_classic() +
  theme(text = element_text(size = 20), 
        axis.text.x = element_text(size = 18), 
        axis.text.y = element_text(size = 18))
### 2-back ###
# vis
f2_2back_vis_tol <- 
  ggplot(data_2back[data_2back$network == "vis",], aes(cbcl_scr_syn_totprob_r, dis)) + 
  geom_hex() + stat_smooth(color="red", fill="red", method=lm, size=0.5, level=0.999) +
  labs(x="total problems score", y="dissimilarities", title="Vis", fill="count") +
  scale_x_continuous(expand = c(0,0), limits = c(-10,180)) +
  scale_y_continuous(expand = c(0,0), limits = c(-0.1,2.1)) +
  theme_classic() +
  theme(text = element_text(size = 20), 
        axis.text.x = element_text(size = 18), 
        axis.text.y = element_text(size = 18))
# lim
f2_2back_lim_tol <- 
  ggplot(data_2back[data_2back$network == "lim",], aes(cbcl_scr_syn_totprob_r, dis)) + 
  geom_hex() + stat_smooth(color="red", fill="red", method=lm, size=0.5, level=0.999) +
  labs(x="total problems score", y="dissimilarities", title="Lim", fill="count") +
  scale_x_continuous(expand = c(0,0), limits = c(-10,180)) +
  scale_y_continuous(expand = c(0,0), limits = c(-0.1,2.1)) +
  theme_classic() +
  theme(text = element_text(size = 20), 
        axis.text.x = element_text(size = 18), 
        axis.text.y = element_text(size = 18))
# v1
f2_2back_v1_tol <- 
  ggplot(data_2back[data_2back$network == "v1",], aes(cbcl_scr_syn_totprob_r, dis)) + 
  geom_hex() + stat_smooth(color="red", fill="red", method=lm, size=0.5, level=0.999) +
  labs(x="total problems score", y="dissimilarities", title="V1", fill="count") +
  scale_x_continuous(expand = c(0,0), limits = c(-10,180)) +
  scale_y_continuous(expand = c(0,0), limits = c(-0.1,2.1)) +
  theme_classic() +
  theme(text = element_text(size = 20), 
        axis.text.x = element_text(size = 18), 
        axis.text.y = element_text(size = 18))

f2_0back_vis_tol
f2_0back_lim_tol
f2_0back_v1_tol
f2_2back_vis_tol
f2_2back_lim_tol
f2_2back_v1_tol
```

(3.4) internalizing score plot
```{r}
# cbcl_scr_syn_internal_r
### 0-back ###
# vis
f2_0back_vis_int <- 
  ggplot(data_0back[data_0back$network == "vis",], aes(cbcl_scr_syn_internal_r, dis)) + 
  geom_hex() + stat_smooth(color="red", fill="red", method=lm, size=0.5, level=0.999) +
  labs(x="internalizing problems score", y="dissimilarities", title="Vis", fill="count") +
  scale_x_continuous(expand = c(0,0), limits = c(-2,60)) +
  scale_y_continuous(expand = c(0,0), limits = c(-0.1,2.1)) +
  theme_classic() +
  theme(text = element_text(size = 20), 
        axis.text.x = element_text(size = 18), 
        axis.text.y = element_text(size = 18))
# lim
f2_0back_lim_int <- 
  ggplot(data_0back[data_0back$network == "lim",], aes(cbcl_scr_syn_internal_r, dis)) + 
  geom_hex() + stat_smooth(color="red", fill="red", method=lm, size=0.5, level=0.999) +
  labs(x="internalizing problems score", y="dissimilarities", title="Lim", fill="count") +
  scale_x_continuous(expand = c(0,0), limits = c(-2,60)) +
  scale_y_continuous(expand = c(0,0), limits = c(-0.1,2.1)) +  
  theme_classic() +
  theme(text = element_text(size = 20), 
        axis.text.x = element_text(size = 18), 
        axis.text.y = element_text(size = 18))
# v1
f2_0back_v1_int <- 
  ggplot(data_0back[data_0back$network == "v1",], aes(cbcl_scr_syn_internal_r, dis)) + 
  geom_hex() + stat_smooth(color="red", fill="red", method=lm, size=0.5, level=0.999) +
  labs(x="internalizing problems score", y="dissimilarities", title="V1", fill="count") +
  scale_x_continuous(expand = c(0,0), limits = c(-2,55)) +
  scale_y_continuous(expand = c(0,0), limits = c(-0.1,2.1)) +  
  theme_classic() +
  theme(text = element_text(size = 20), 
        axis.text.x = element_text(size = 18), 
        axis.text.y = element_text(size = 18))
### 2-back ###
# vis
f2_2back_vis_int <- 
  ggplot(data_2back[data_2back$network == "vis",], aes(cbcl_scr_syn_internal_r, dis)) + 
  geom_hex() + stat_smooth(color="red", fill="red", method=lm, size=0.5, level=0.999) +
  labs(x="internalizing problems score", y="dissimilarities", title="Vis", fill="count") +
  scale_x_continuous(expand = c(0,0), limits = c(-2,55)) +
  scale_y_continuous(expand = c(0,0), limits = c(-0.1,2.1)) +  
  theme_classic() +
  theme(text = element_text(size = 20), 
        axis.text.x = element_text(size = 18), 
        axis.text.y = element_text(size = 18))
# lim
f2_2back_lim_int <- 
  ggplot(data_2back[data_2back$network == "lim",], aes(cbcl_scr_syn_internal_r, dis)) + 
  geom_hex() + stat_smooth(color="red", fill="red", method=lm, size=0.5, level=0.999) +
  labs(x="internalizing problems score", y="dissimilarities", title="Lim", fill="count") +
  scale_x_continuous(expand = c(0,0), limits = c(-2,55)) +
  scale_y_continuous(expand = c(0,0), limits = c(-0.1,2.1)) +  
  theme_classic() +
  theme(text = element_text(size = 20), 
        axis.text.x = element_text(size = 18), 
        axis.text.y = element_text(size = 18))
# v1
f2_2back_v1_int <- 
  ggplot(data_2back[data_2back$network == "v1",], aes(cbcl_scr_syn_internal_r, dis)) + 
  geom_hex() + stat_smooth(color="red", fill="red", method=lm, size=0.5, level=0.999) +
  labs(x="internalizing problems score", y="dissimilarities", title="V1", fill="count") +
  scale_x_continuous(expand = c(0,0), limits = c(-2,55)) +
  scale_y_continuous(expand = c(0,0), limits = c(-0.1,2.1)) +  
  theme_classic() +
  theme(text = element_text(size = 20), 
        axis.text.x = element_text(size = 18), 
        axis.text.y = element_text(size = 18))

f2_0back_vis_int
f2_0back_lim_int
f2_0back_v1_int
f2_2back_vis_int
f2_2back_lim_int
f2_2back_v1_int
```

(3.5) externalizing score plot
```{r}
# cbcl_scr_syn_external_r
### 0-back ###
# vis
f2_0back_vis_ext <- 
  ggplot(data_0back[data_0back$network == "vis",], aes(cbcl_scr_syn_external_r, dis)) + 
  geom_hex() + stat_smooth(color="red", fill="red", method=lm, size=0.5, level=0.999) +
  labs(x="externalizing problems score", y="dissimilarities", title="Vis", fill="count") +
  scale_x_continuous(expand = c(0,0), limits = c(-2,55)) +
  scale_y_continuous(expand = c(0,0), limits = c(-0.1,2.1)) + 
  theme_classic() +
  theme(text = element_text(size = 20), 
        axis.text.x = element_text(size = 18), 
        axis.text.y = element_text(size = 18))
# lim
f2_0back_lim_ext <- 
  ggplot(data_0back[data_0back$network == "lim",], aes(cbcl_scr_syn_external_r, dis)) + 
  geom_hex() + stat_smooth(color="red", fill="red", method=lm, size=0.5, level=0.999) +
  labs(x="externalizing problems score", y="dissimilarities", title="Lim", fill="count") +
  scale_x_continuous(expand = c(0,0), limits = c(-2,55)) +
  scale_y_continuous(expand = c(0,0), limits = c(-0.1,2.1)) + 
  theme_classic() +
  theme(text = element_text(size = 20), 
        axis.text.x = element_text(size = 18), 
        axis.text.y = element_text(size = 18))
# v1
f2_0back_v1_ext <- 
  ggplot(data_0back[data_0back$network == "v1",], aes(cbcl_scr_syn_external_r, dis)) + 
  geom_hex() + stat_smooth(color="red", fill="red", method=lm, size=0.5, level=0.999) +
  labs(x="externalizing problems score", y="dissimilarities", title="V1", fill="count") +
  scale_x_continuous(expand = c(0,0), limits = c(-2,55)) +
  scale_y_continuous(expand = c(0,0), limits = c(-0.1,2.1)) + 
  theme_classic() +
  theme(text = element_text(size = 20), 
        axis.text.x = element_text(size = 18), 
        axis.text.y = element_text(size = 18))
### 2-back ###
# vis
f2_2back_vis_ext <- 
  ggplot(data_2back[data_2back$network == "vis",], aes(cbcl_scr_syn_external_r, dis)) + 
  geom_hex() + stat_smooth(color="red", fill="red", method=lm, size=0.5, level=0.999) +
  labs(x="externalizing problems score", y="dissimilarities", title="Vis", fill="count") +
  scale_x_continuous(expand = c(0,0), limits = c(-2,55)) +
  scale_y_continuous(expand = c(0,0), limits = c(-0.1,2.1)) + 
  theme_classic() +
  theme(text = element_text(size = 20), 
        axis.text.x = element_text(size = 18), 
        axis.text.y = element_text(size = 18))
# lim
f2_2back_lim_ext <- 
  ggplot(data_2back[data_2back$network == "lim",], aes(cbcl_scr_syn_external_r, dis)) + 
  geom_hex() + stat_smooth(color="red", fill="red", method=lm, size=0.5, level=0.999) +
  labs(x="externalizing problems score", y="dissimilarities", title="Lim", fill="count") +
  scale_x_continuous(expand = c(0,0), limits = c(-2,55)) +
  scale_y_continuous(expand = c(0,0), limits = c(-0.1,2.1)) + 
  theme_classic() +
  theme(text = element_text(size = 20), 
        axis.text.x = element_text(size = 18), 
        axis.text.y = element_text(size = 18))
# v1
f2_2back_v1_ext <- 
  ggplot(data_2back[data_2back$network == "v1",], aes(cbcl_scr_syn_external_r, dis)) + 
  geom_hex() + stat_smooth(color="red", fill="red", method=lm, size=0.5, level=0.999) +
  labs(x="externalizing problems score", y="dissimilarities", title="V1", fill="count") +
  scale_x_continuous(expand = c(0,0), limits = c(-2,55)) +
  scale_y_continuous(expand = c(0,0), limits = c(-0.1,2.1)) + 
  theme_classic() +
  theme(text = element_text(size = 20), 
        axis.text.x = element_text(size = 18), 
        axis.text.y = element_text(size = 18))

f2_0back_vis_ext
f2_0back_lim_ext
f2_0back_v1_ext
f2_2back_vis_ext
f2_2back_lim_ext
f2_2back_v1_ext
```

# part 3 - dissimilarity and performance
- (1) process datasets
```{r}
# Calculate the mean accuracy of all faces
# 0-back
data_0back <- data_0back %>% 
  mutate(data_0back, allface_0b_rate = rowMeans(select(data_0back, c("tfmri_nb_all_beh_c0bpf_rate","tfmri_nb_all_beh_c0bnf_rate", "tfmri_nb_all_beh_c0bngf_rate"))))

# 2-back 
data_2back <- data_2back %>%
  mutate(data_2back, allface_2b_rate = rowMeans(select(data_2back, c("tfmri_nb_all_beh_c2bpf_rate","tfmri_nb_all_beh_c2bnf_rate","tfmri_nb_all_beh_c2bngf_rate"))))

# head(data_0back)
# head(data_2back)
```

- (2) mixed models
```{r}
###### controlled for overall performance ######
## 0-back ##
# vis #
lmer3_0back_vis <- lmer(allface_0b_rate ~ dis + tfmri_nb_all_beh_c0b_rate + sex + interview_age + (1|site_id_l/rel_family_id), data = data_0back[data_0back$network == "vis",])
summary(lmer3_0back_vis) # 0.8330

# lim #
lmer3_0back_lim <- lmer(allface_0b_rate ~ dis + tfmri_nb_all_beh_c0b_rate + sex + interview_age + (1|site_id_l/rel_family_id), data = data_0back[data_0back$network == "lim",])
summary(lmer3_0back_lim) # 0.1511

# v1
lmer3_0back_v1 <- lmer(allface_0b_rate ~ dis + tfmri_nb_all_beh_c0b_rate + sex + interview_age + (1|site_id_l/rel_family_id), data = data_0back[data_0back$network == "v1",])
summary(lmer3_0back_v1) # 0.391

## 2-back ##
# vis #
lmer3_2back_vis <- lmer(allface_2b_rate ~ dis + tfmri_nb_all_beh_c2b_rate + sex + interview_age + (1|site_id_l/rel_family_id), data = data_2back[data_2back$network == "vis",])
summary(lmer3_2back_vis) # 0.000977 ***

# lim
lmer3_2back_lim <- lmer(allface_2b_rate ~ dis + tfmri_nb_all_beh_c2b_rate + sex + interview_age + (1|site_id_l/rel_family_id), data = data_2back[data_2back$network == "lim",])
summary(lmer3_2back_lim) # 0.00492 **

# v1
lmer3_2back_v1 <- lmer(allface_2b_rate ~ dis + tfmri_nb_all_beh_c2b_rate + sex + interview_age + (1|site_id_l/rel_family_id), data = data_2back[data_2back$network == "v1",])
summary(lmer3_2back_v1) # 0.0215 *
```

- (3) plots
```{r}
# scatter plot of all three dissimilarity and cbcl scores
## 0-back
# vis
f3_0back_vis <- 
  ggplot(data_0back[data_0back$network == "vis",], aes(x=dis, y=allface_0b_rate)) + 
  geom_hex() + stat_smooth(color="red", fill="red", method=lm, size=0.5, level=0.999) +
  labs(title = "Vis (0-back)",
       x = "Dissimilarity",
       y = "Accuracy") +
  scale_x_continuous(expand = c(0,0), limits = c(0,2)) +
  scale_y_continuous(expand = c(0,0), limits = c(0.5,1.1)) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"), text = element_text(size = 18))
# lim
f3_0back_lim <- 
  ggplot(data_0back[data_0back$network == "lim",], aes(x=dis, y=allface_0b_rate)) + 
  geom_hex() + stat_smooth(color="red", fill="red", method=lm, size=0.5, level=0.999) +
  labs(title = "Lim (0-back)",
       x = "Dissimilarity",
       y = "Accuracy") +
  scale_x_continuous(expand = c(0,0), limits = c(0,2)) +
  scale_y_continuous(expand = c(0,0), limits = c(0.5,1.1)) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"), text = element_text(size = 18))
# v1
f3_0back_v1 <- 
  ggplot(data_0back[data_0back$network == "v1",], aes(x=dis, y=allface_0b_rate)) + 
  geom_hex() + stat_smooth(color="red", fill="red", method=lm, size=0.5, level=0.999) +
  labs(title = "V1 (0-back)",
       x = "Dissimilarity",
       y = "Accuracy") +
  scale_x_continuous(expand = c(0,0), limits = c(0,2)) +
  scale_y_continuous(expand = c(0,0), limits = c(0.5,1.1)) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"), text = element_text(size = 18))

## 2-back
# vis
f3_2back_vis <- 
  ggplot(data_2back[data_2back$network == "vis",], aes(x=dis, y=allface_2b_rate)) + 
  geom_hex() + stat_smooth(color="red", fill="red", method=lm, size=0.5, level=0.999) +
  labs(title = "Vis (2-back)",
       x = "Dissimilarity",
       y = "Accuracy") +
  scale_x_continuous(expand = c(0,0), limits = c(0,2)) +
  scale_y_continuous(expand = c(0,0), limits = c(0.5,1.1)) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"), text = element_text(size = 18))
# lim
f3_2back_lim <- 
  ggplot(data_2back[data_2back$network == "lim",], aes(x=dis, y=allface_2b_rate)) + 
  geom_hex() + stat_smooth(color="red", fill="red", method=lm, size=0.5, level=0.999) +
  labs(title = "Lim (2-back)",
       x = "Dissimilarity",
       y = "Accuracy") +
  scale_x_continuous(expand = c(0,0), limits = c(0,2)) +
  scale_y_continuous(expand = c(0,0), limits = c(0.5,1.1)) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"), text = element_text(size = 18))
# v1
f3_2back_v1 <- 
  ggplot(data_2back[data_2back$network == "v1",], aes(x=dis, y=allface_2b_rate)) + 
  geom_hex() + stat_smooth(color="red", fill="red", method=lm, size=0.5, level=0.999) +
  labs(title = "V1 (2-back)",
       x = "Dissimilarity",
       y = "Accuracy") +
  scale_x_continuous(expand = c(0,0), limits = c(0,2)) +
  scale_y_continuous(expand = c(0,0), limits = c(0.5,1.1)) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"), text = element_text(size = 18))

f3_0back_vis
f3_0back_lim
f3_0back_v1
f3_2back_vis
f3_2back_lim
f3_2back_v1
```

# part 4 - effect sizes
- (1) process dataset for effect sizes
```{r}
# merge the dataframes
data_hf_0b_eff <- merge(data_hf_0b, fam_site_id, by="src_subject_id")
data_hn_0b_eff <- merge(data_hn_0b, fam_site_id, by="src_subject_id")
data_fn_0b_eff <- merge(data_fn_0b, fam_site_id, by="src_subject_id")
data_hf_2b_eff <- merge(data_hf_2b, fam_site_id, by="src_subject_id")
data_hn_2b_eff <- merge(data_hn_2b, fam_site_id, by="src_subject_id")
data_fn_2b_eff <- merge(data_fn_2b, fam_site_id, by="src_subject_id")

data_hf_0b_eff$nback <- "0_back"
data_hn_0b_eff$nback <- "0_back"
data_fn_0b_eff$nback <- "0_back"
data_hf_2b_eff$nback <- "2_back"
data_hn_2b_eff$nback <- "2_back"
data_fn_2b_eff$nback <- "2_back"

data_hf_0b_eff$condition <- "hf"
data_hn_0b_eff$condition <- "hn"
data_fn_0b_eff$condition <- "fn"
data_hf_2b_eff$condition <- "hf"
data_hn_2b_eff$condition <- "hn"
data_fn_2b_eff$condition <- "fn"

data_hf_0b_eff$v1 <- NULL
data_hn_0b_eff$v1 <- NULL
data_fn_0b_eff$v1 <- NULL
data_hf_2b_eff$v1 <- NULL
data_hn_2b_eff$v1 <- NULL
data_fn_2b_eff$v1 <- NULL

# (6.1) sanity check
dim(data_hf_0b_eff) # 4952    14
dim(data_hf_2b_eff) # 4952    14
dim(data_hn_0b_eff) # 4952    14
dim(data_hn_2b_eff) # 4952    14
dim(data_fn_0b_eff) # 4952    14
dim(data_fn_2b_eff) # 4952    14

# head(data_hf_0b_eff)
```

- (2) Process dataframe for mixed effect models
one model for each network - association between 1-r and each network (for example, vis vs. non-vis)
```{r}
list_eff <- c("vis","lim","dor","ven","fro","def","som")

### 0-back ###
# convert the data into long format
data_hf_0b_long <- pivot_longer(data_hf_0b_eff, cols = all_of(list_eff), names_to = "network", values_to = "dis")
data_hn_0b_long <- pivot_longer(data_hn_0b_eff, cols = all_of(list_eff), names_to = "network", values_to = "dis")
data_fn_0b_long <- pivot_longer(data_fn_0b_eff, cols = all_of(list_eff), names_to = "network", values_to = "dis")

# COMBINE ALL THREE CONDITIONS 
data_0b <- rbind(data_hf_0b_long, data_hn_0b_long, data_fn_0b_long)
# convert vector objects to factors
data_0b <- vector_to_factor(data_0b)

# setup the data for mixed effect model
data_0b$vis <- (data_0b$network == "vis") * 1
data_0b$lim <- (data_0b$network == "lim") * 1
data_0b$dor <- (data_0b$network == "dor") * 1
data_0b$ven <- (data_0b$network == "ven") * 1
data_0b$fro <- (data_0b$network == "fro") * 1
data_0b$def <- (data_0b$network == "def") * 1
data_0b$som <- (data_0b$network == "som") * 1

data_0b$vis <- as.factor(data_0b$vis)
data_0b$lim <- as.factor(data_0b$lim)
data_0b$dor <- as.factor(data_0b$dor)
data_0b$ven <- as.factor(data_0b$ven)
data_0b$fro <- as.factor(data_0b$fro)
data_0b$def <- as.factor(data_0b$def)
data_0b$som <- as.factor(data_0b$som)

### 2-back ###
# convert the data into long format
data_hf_2b_long <- pivot_longer(data_hf_2b_eff, cols = all_of(list_eff), names_to = "network", values_to = "dis")
data_hn_2b_long <- pivot_longer(data_hn_2b_eff, cols = all_of(list_eff), names_to = "network", values_to = "dis")
data_fn_2b_long <- pivot_longer(data_fn_2b_eff, cols = all_of(list_eff), names_to = "network", values_to = "dis")

# COMBINE ALL THREE CONDITIONS 
data_2b <- rbind(data_hf_2b_long, data_hn_2b_long, data_fn_2b_long)
# convert vector objects to factors
data_2b <- vector_to_factor(data_2b)

# setup the data for mixed effect model
data_2b$vis <- (data_2b$network == "vis") * 1
data_2b$lim <- (data_2b$network == "lim") * 1
data_2b$dor <- (data_2b$network == "dor") * 1
data_2b$ven <- (data_2b$network == "ven") * 1
data_2b$fro <- (data_2b$network == "fro") * 1
data_2b$def <- (data_2b$network == "def") * 1
data_2b$som <- (data_2b$network == "som") * 1

data_2b$vis <- as.factor(data_2b$vis)
data_2b$lim <- as.factor(data_2b$lim)
data_2b$dor <- as.factor(data_2b$dor)
data_2b$ven <- as.factor(data_2b$ven)
data_2b$fro <- as.factor(data_2b$fro)
data_2b$def <- as.factor(data_2b$def)
data_2b$som <- as.factor(data_2b$som)
```

- (3) run mixed models
use R marginal squared as effect sizes of each network
```{r}
### 0-back ###
# run mixed effect model for seven networks separately
lme1_vis_0b <- lmer(dis ~ vis * condition + (1|src_subject_id) + (1|site_id_l/rel_family_id), data = data_0b)
lme1_lim_0b <- lmer(dis ~ lim * condition + (1|src_subject_id) + (1|site_id_l/rel_family_id), data = data_0b)
lme1_dor_0b <- lmer(dis ~ dor * condition + (1|src_subject_id) + (1|site_id_l/rel_family_id), data = data_0b)
lme1_ven_0b <- lmer(dis ~ ven * condition + (1|src_subject_id) + (1|site_id_l/rel_family_id), data = data_0b)
lme1_fro_0b <- lmer(dis ~ fro * condition + (1|src_subject_id) + (1|site_id_l/rel_family_id), data = data_0b)
lme1_def_0b <- lmer(dis ~ def * condition + (1|src_subject_id) + (1|site_id_l/rel_family_id), data = data_0b)
lme1_som_0b <- lmer(dis ~ som * condition + (1|src_subject_id) + (1|site_id_l/rel_family_id), data = data_0b)

# get the R squared marginal values
m <- matrix(ncol = 0, nrow = 0) # create an empty matrix
r_squared_0back <- data.frame(m) # empty matrix to dataframe
r_squared_0back <- r.squaredGLMM(lme1_vis_0b) %>% 
  rbind(r_squared_0back, r.squaredGLMM(lme1_lim_0b)) %>%
  rbind(r_squared_0back, r.squaredGLMM(lme1_dor_0b)) %>%
  rbind(r_squared_0back, r.squaredGLMM(lme1_ven_0b)) %>%
  rbind(r_squared_0back, r.squaredGLMM(lme1_fro_0b)) %>%
  rbind(r_squared_0back, r.squaredGLMM(lme1_def_0b)) %>%
  rbind(r_squared_0back, r.squaredGLMM(lme1_som_0b))
row.names(r_squared_0back) <- list_eff

### 2-back ###
# run mixed effect model for seven networks separately
lme1_vis_2b <- lmer(dis ~ vis * condition + (1|src_subject_id) + (1|site_id_l/rel_family_id), data = data_2b)
lme1_lim_2b <- lmer(dis ~ lim * condition + (1|src_subject_id) + (1|site_id_l/rel_family_id), data = data_2b)
lme1_dor_2b <- lmer(dis ~ dor * condition + (1|src_subject_id) + (1|site_id_l/rel_family_id), data = data_2b)
lme1_ven_2b <- lmer(dis ~ ven * condition + (1|src_subject_id) + (1|site_id_l/rel_family_id), data = data_2b)
lme1_fro_2b <- lmer(dis ~ fro * condition + (1|src_subject_id) + (1|site_id_l/rel_family_id), data = data_2b)
lme1_def_2b <- lmer(dis ~ def * condition + (1|src_subject_id) + (1|site_id_l/rel_family_id), data = data_2b)
lme1_som_2b <- lmer(dis ~ som * condition + (1|src_subject_id) + (1|site_id_l/rel_family_id), data = data_2b)

# get the R squared marginal values
m <- matrix(ncol = 0, nrow = 0) # create an empty matrix
r_squared_2back <- data.frame(m) # empty matrix to dataframe
r_squared_2back <- r.squaredGLMM(lme1_vis_2b) %>% 
  rbind(r_squared_2back, r.squaredGLMM(lme1_lim_2b)) %>%
  rbind(r_squared_2back, r.squaredGLMM(lme1_dor_2b)) %>%
  rbind(r_squared_2back, r.squaredGLMM(lme1_ven_2b)) %>%
  rbind(r_squared_2back, r.squaredGLMM(lme1_fro_2b)) %>%
  rbind(r_squared_2back, r.squaredGLMM(lme1_def_2b)) %>%
  rbind(r_squared_2back, r.squaredGLMM(lme1_som_2b))
row.names(r_squared_2back) <- list_eff

r_squared_0back
r_squared_2back
```

- (4) plot - effect sizes bar plots
```{r}
df_plot = data.frame(network=rownames(r_squared_0back),
                     r2m_0back=as.numeric(r_squared_0back[-1][,1]),
                     r2m_2back=as.numeric(r_squared_2back[-1][,1]))

# list_eff <- c("vis","lim","dor","ven","fro","def","som") # defined before
plot_order_of_networks <- c("dor","ven","fro","lim","def","vis","som")

# 0-back
f1_0back <- ggplot(data=df_plot, aes(x = network, y = r2m_0back)) + 
  geom_bar(stat = "identity", position = "dodge", color = "black", fill = "white") +
  aes(x = factor(network, plot_order_of_networks), y = r2m_0back) +
  labs(x = "Functional networks",
       y = "Effect sizes") +
  scale_y_continuous(expand = c(0,0), limits = c(0,0.6)) +
  labs(title = "0-back") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"), text = element_text(size = 20), axis.text.x = element_text(angle = 90, hjust = 1), aspect.ratio = 1)
f1_0back

# 2-back
f1_2back <- ggplot(data=df_plot, aes(x = network, y = r2m_2back)) + 
  geom_bar(stat = "identity", position = "dodge", color = "black", fill = "white") +
  aes(x = factor(network, plot_order_of_networks), y = r2m_2back) +
  labs(x = "Functional networks",
       y = "Effect sizes") +
  scale_y_continuous(expand = c(0,0), limits = c(0,0.6)) +
  labs(title = "2-back") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"), text = element_text(size = 20), axis.text.x = element_text(angle = 90, hjust = 1), aspect.ratio = 1)
f1_2back
```

- (5) cohen.d
```{r}
# 0-back
cohen.d(data_0back$dis[data_0back$network == "vis"], data_0back$dis[!(data_0back$network == "vis")]) # -0.7819427 (medium)
cohen.d(data_0back$dis[data_0back$network == "lim"], data_0back$dis[!(data_0back$network == "lim")]) #  1.161041 (large)
cohen.d(data_0back$dis[data_0back$network == "fro"], data_0back$dis[!(data_0back$network == "fro")]) #  0.1922416 (negligible)
cohen.d(data_0back$dis[data_0back$network == "dor"], data_0back$dis[!(data_0back$network == "dor")]) # -0.1743875 (negligible)
cohen.d(data_0back$dis[data_0back$network == "ven"], data_0back$dis[!(data_0back$network == "ven")]) #  0.2764237 (small)
cohen.d(data_0back$dis[data_0back$network == "def"], data_0back$dis[!(data_0back$network == "def")]) #  0.4370847 (small)
cohen.d(data_0back$dis[data_0back$network == "som"], data_0back$dis[!(data_0back$network == "som")]) #  0.4547063 (small)
# 2-back
cohen.d(data_2back$dis[data_2back$network == "vis"], data_2back$dis[!(data_2back$network == "vis")]) # -0.5052931 (medium)
cohen.d(data_2back$dis[data_2back$network == "lim"], data_2back$dis[!(data_2back$network == "lim")]) #  1.417457 (large)
cohen.d(data_2back$dis[data_2back$network == "fro"], data_2back$dis[!(data_2back$network == "fro")]) # -0.1085989 (negligible)
cohen.d(data_2back$dis[data_2back$network == "dor"], data_2back$dis[!(data_2back$network == "dor")]) # -0.3534341 (small)
cohen.d(data_2back$dis[data_2back$network == "ven"], data_2back$dis[!(data_2back$network == "ven")]) #  0.01807474 (negligible)
cohen.d(data_2back$dis[data_2back$network == "def"], data_2back$dis[!(data_2back$network == "def")]) #  0.2594864 (small)
cohen.d(data_2back$dis[data_2back$network == "som"], data_2back$dis[!(data_2back$network == "som")]) #  0.527744 (medium)
```

# part 5 - performance and total problem score
- (1) data processing
```{r}
# for three way interactions
# preprocess subj_list
names(subj_list)[names(subj_list) == 'V1'] <- 'src_subject_id'
subj_list$src_subject_id <- gsub("NDAR_", "", subj_list$src_subject_id)

# merge dataframes
nback_sub <- nback[,c("src_subject_id","tfmri_nb_all_beh_c0bpf_rate","tfmri_nb_all_beh_c0bnf_rate","tfmri_nb_all_beh_c0bngf_rate","tfmri_nb_all_beh_c2bpf_rate","tfmri_nb_all_beh_c2bnf_rate","tfmri_nb_all_beh_c2bngf_rate")]
perf_psy <- merge(fam_site_id, nback_sub, by="src_subject_id")
perf_psy <- merge(perf_psy, abcd_cbcl, by="src_subject_id")
perf_psy <- merge(perf_psy, subj_list, by="src_subject_id")
perf_psy <- perf_psy %>%
  mutate_at(c("interview_age","cbcl_scr_syn_totprob_t","cbcl_scr_syn_totprob_r","tfmri_nb_all_beh_c0bpf_rate","tfmri_nb_all_beh_c0bnf_rate","tfmri_nb_all_beh_c0bngf_rate","tfmri_nb_all_beh_c2bpf_rate","tfmri_nb_all_beh_c2bnf_rate","tfmri_nb_all_beh_c2bngf_rate"), as.numeric)
# head(perf_psy)
dim(perf_psy) # 5003   21

# select subjects with clean imaging data
perf_psy <- merge(data_hf_0b["src_subject_id"], perf_psy, by="src_subject_id")
# head(perf_psy)
dim(perf_psy) # 4952   21

# pivot longer 
perf_psy_long <- 
  pivot_longer(perf_psy, cols = c("tfmri_nb_all_beh_c0bpf_rate","tfmri_nb_all_beh_c0bnf_rate","tfmri_nb_all_beh_c0bngf_rate","tfmri_nb_all_beh_c2bpf_rate","tfmri_nb_all_beh_c2bnf_rate","tfmri_nb_all_beh_c2bngf_rate"), names_to = "nback_condition", values_to = "nback_perf")

# extract 0-back and 2-back & emotions 
perf_psy_long$nback_condition <- gsub("tfmri_nb_all_beh_","",as.character(perf_psy_long$nback_condition))
perf_psy_long$nback_condition <- gsub("_rate","",as.character(perf_psy_long$nback_condition))
perf_psy_long$nback_condition <- gsub("b","_b",as.character(perf_psy_long$nback_condition))
perf_psy_long <- perf_psy_long %>% separate(nback_condition,c("nback","emotion"))
perf_psy_long <- perf_psy_long %>%
  mutate_at(c("nback","emotion"), as.factor)
# head(perf_psy_long)
dim(perf_psy_long) # 30018    18
```

- (2) full mix model
```{r}
lmer6_acc_psych_threelevel <- lmer(nback_perf ~ cbcl_scr_syn_totprob_r * nback * emotion + sex + interview_age + (1|site_id_l/rel_family_id), data = perf_psy_long)
summary(lmer6_acc_psych_threelevel)
tab_model(lmer6_acc_psych_threelevel, digits = 4)
```

- (3) three-way anova
```{r}
anova_model1 <- aov(nback_perf ~ cbcl_scr_syn_totprob_r * nback * emotion, data = perf_psy_long)
summary(anova_model1)
tab_model(anova_model1)
```

- (4) plot
```{r}
# Create the scatter plot and add the regression line
p <- ggplot(data = perf_psy_long, aes(x=cbcl_scr_syn_totprob_r, y=nback_perf, color=emotion, shape=emotion)) +
  geom_point(aes(color=emotion),alpha = 0.5) +
  facet_wrap(~nback) +
  geom_smooth(method = "lm", se = TRUE, alpha = 0.1) +
  scale_color_manual(values=c('#999999','#56B4E9','#E69F00')) +
  labs(title = "", x = "CBCL total problems score", y = "Accuracy") +
  theme_classic() +
  theme(text = element_text(size = 20))
p
```

# part 6 - demographics table
- (1) preprocessing
```{r}
# demographic table includes 
# (1) age
# (2) sex assigned at birth; female (%)
# (3) race/ethnicity (%)
# (4) family income (%)
# (5) caregiver education (%) 

# select subjects with clean imaging data
abcd_pdem <- merge(perf_psy["src_subject_id"], abcd_pdem, by="src_subject_id")
abcd_pdem <- merge(abcd_pdem, fam_site_id, by="src_subject_id")
abcd_pdem <- abcd_pdem %>%
  mutate_at("interview_age", as.numeric) 

# sanity check
dim(abcd_pdem) # 4952   47
# head(abcd_pdem)

# re-code household income into three levels
household.income = abcd_pdem$demo_comb_income_v2
household.income[abcd_pdem$demo_comb_income_v2 == "1"] = 1 # "[<50K]"
household.income[abcd_pdem$demo_comb_income_v2 == "2"] = 1 # "[<50K]"
household.income[abcd_pdem$demo_comb_income_v2 == "3"] = 1 # "[<50K]"
household.income[abcd_pdem$demo_comb_income_v2 == "4"] = 1 # "[<50K]"
household.income[abcd_pdem$demo_comb_income_v2 == "5"] = 1 # "[<50K]"
household.income[abcd_pdem$demo_comb_income_v2 == "6"] = 1 # "[<50K]"
household.income[abcd_pdem$demo_comb_income_v2 == "7"] = 2 # "[>=50K & <100K]"
household.income[abcd_pdem$demo_comb_income_v2 == "8"] = 2 # "[>=50K & <100K]"
household.income[abcd_pdem$demo_comb_income_v2 == "9"] = 3 # "[>=100K]"
household.income[abcd_pdem$demo_comb_income_v2 == "10"] = 3 # "[>=100K]"
household.income[abcd_pdem$demo_comb_income_v2 == "777"] = NA
household.income[abcd_pdem$demo_comb_income_v2 == "999"] = NA
household.income[household.income %in% c(NA, "999", "777")] = NA
abcd_pdem$household.income = factor(household.income, levels= 1:3, labels = c("[<50K]", "[>=50K & <100K]", "[>=100K]") )

# re-code caregiver education into five levels
high.educ1 = abcd_pdem$demo_prnt_ed_v2
high.educ2 = abcd_pdem$demo_prtnr_ed_v2
high.educ1[which(high.educ1 == "999")] = NA
high.educ2[which(high.educ2 == "999")] = NA
high.educ1[which(high.educ1 == "777")] = NA
high.educ2[which(high.educ2 == "777")] = NA
high.educ = pmax(as.numeric(as.character(high.educ1)), as.numeric(as.character(high.educ2)), na.rm=T)
idx <- which(high.educ %in% 0:12, arr.ind = TRUE)
high.educ[idx] = 1 # "< HS Diploma"
idx <- which(high.educ %in% 13:14, arr.ind = TRUE)
high.educ[idx] = 2 # "HS Diploma/GED"
idx <- which(high.educ %in% 15:17, arr.ind = TRUE)
high.educ[idx] = 3 # "Some College"
idx <- which(high.educ == 18, arr.ind = TRUE)
high.educ[idx] = 4 # "Bachelor"
idx <- which(high.educ %in% 19:21, arr.ind = TRUE)
high.educ[idx] = 5 # "Post Graduate Degree"
high.educ[which(high.educ == "999")]=NA
high.educ[which(high.educ == "777")]=NA
abcd_pdem$high.educ = factor(high.educ, levels= 1:5, labels = c("< HS Diploma","HS Diploma/GED","Some College","Bachelor","Post Graduate Degree") )

# re-code racial categories with more easily identifiable variable names (into 8 levels)
abcd_pdem$white= (abcd_pdem$demo_race_a_p___10 == 1)*1
abcd_pdem$black= (abcd_pdem$demo_race_a_p___11 == 1)*1
abcd_pdem$asian = 0
abcd_pdem$asian[abcd_pdem$demo_race_a_p___18 == 1 | abcd_pdem$demo_race_a_p___19 == 1 | 
                abcd_pdem$demo_race_a_p___20 == 1 | abcd_pdem$demo_race_a_p___21 == 1 | 
                abcd_pdem$demo_race_a_p___22 == 1 | abcd_pdem$demo_race_a_p___23 == 1 |
                abcd_pdem$demo_race_a_p___24==1] = 1
abcd_pdem$aian = 0
abcd_pdem$aian[abcd_pdem$demo_race_a_p___12 == 1 | abcd_pdem$demo_race_a_p___13 == 1] = 1
abcd_pdem$nhpi = 0
abcd_pdem$nhpi[abcd_pdem$demo_race_a_p___14 == 1 | abcd_pdem$demo_race_a_p___15 == 1 | 
               abcd_pdem$demo_race_a_p___16 == 1 | abcd_pdem$demo_race_a_p___17 == 1] = 1
abcd_pdem$other = 0
abcd_pdem$other[abcd_pdem$demo_race_a_p___25 == 1] = 1

# generate mixed category
abcd_pdem$mixed = 
  abcd_pdem$white + abcd_pdem$black + abcd_pdem$asian + abcd_pdem$aian + abcd_pdem$nhpi + abcd_pdem$other
abcd_pdem$mixed[abcd_pdem$mixed <= 1] =  0
abcd_pdem$mixed[abcd_pdem$mixed > 1]  =  1

# generate single race.ethnicity variable 
# note: If you want fewer categories, change code here so that different categories (e.g., other, mixed) have same value (e.g., both = 7)
abcd_pdem$race.eth = NA
abcd_pdem$race.eth[abcd_pdem$demo_ethn_v2 == 1] = 1 
abcd_pdem$race.eth[abcd_pdem$white == 1] = 2
abcd_pdem$race.eth[abcd_pdem$black == 1] = 3
abcd_pdem$race.eth[abcd_pdem$asian == 1] = 4
abcd_pdem$race.eth[abcd_pdem$aian  == 1] = 5
abcd_pdem$race.eth[abcd_pdem$nhpi  == 1] = 6
abcd_pdem$race.eth[abcd_pdem$other == 1] = 7
abcd_pdem$race.eth[abcd_pdem$mixed == 1] = 8
abcd_pdem$race.eth <- factor(abcd_pdem$race.eth, levels = 1:8, 
                             labels = c("Hispanic", "White", "Black", "Asian", "AIAN", "NHPI", "Other", "Mixed"))
# head(abcd_pdem)
```

- (2) generate the table
```{r}
# create a summary/demographics table
vars1 <- c("interview_age", "sex", "race.eth", "high.educ", "household.income")
tab1 <- CreateTableOne(vars = vars1, data = abcd_pdem)
tabAsStringMatrix <- print(tab1, printToggle = FALSE, noSpaces = TRUE)
tab1 = knitr::kable(tabAsStringMatrix)
tab1
```

# part 7 - mediation
- (1) mediation analysis - all three emotional comparisons
```{r}
library(mediation)
detach("package:lmerTest", unload = TRUE) # must unload lmerTest to run mediation
# detach("package:psych", unload = TRUE) # also be sure psych is detached

# using average accuracy of all faces
# IV - acc, DV - psy, med - sqED
### 0-back ###
# vis
lme4_0back_vis.0 <- 
  lmer(cbcl_scr_syn_totprob_r ~ allface_0b_rate + sex + interview_age + (1|site_id_l), data = data_0back[data_0back$network == "vis",])
lme4_0back_vis.M <- 
  lmer(dis ~ allface_0b_rate + sex + interview_age + (1|site_id_l), data = data_0back[data_0back$network == "vis",])
lme4_0back_vis.Y <- 
  lmer(cbcl_scr_syn_totprob_r ~ allface_0b_rate + dis + sex + interview_age + (1|site_id_l), data = data_0back[data_0back$network == "vis",])
lme4_med_0back_vis <- 
  mediate(model.m = lme4_0back_vis.M, model.y = lme4_0back_vis.Y, treat='allface_0b_rate', mediator = 'dis', data=data_0back[data_0back$network == "vis",], sims=1000, dropobs=TRUE)

summary(lme4_0back_vis.0)
summary(lme4_0back_vis.M)
summary(lme4_0back_vis.Y)
summary(lme4_med_0back_vis) # ACME            -0.65252     -1.11414        -0.17   0.006 **

# lim
lme4_0back_lim.0 <- 
  lmer(cbcl_scr_syn_totprob_r ~ allface_0b_rate + sex + interview_age + (1|site_id_l), data = data_0back[data_0back$network == "lim",])
lme4_0back_lim.M <- 
  lmer(dis ~ allface_0b_rate + sex + interview_age + (1|site_id_l), data = data_0back[data_0back$network == "lim",])
lme4_0back_lim.Y <- 
  lmer(cbcl_scr_syn_totprob_r ~ allface_0b_rate + dis + sex + interview_age + (1|site_id_l), data = data_0back[data_0back$network == "lim",])
lme4_med_0back_lim <- 
  mediate(model.m = lme4_0back_lim.M, model.y = lme4_0back_lim.Y, treat='allface_0b_rate', mediator = 'dis', data=data_0back[data_0back$network == "lim",], sims=1000, dropobs=TRUE)

summary(lme4_0back_lim.0)
summary(lme4_0back_lim.M)
summary(lme4_0back_lim.Y)
summary(lme4_med_0back_lim) # ACME            -0.08545     -0.26251         0.06    0.26   

# v1
lme4_0back_v1.0 <- 
  lmer(cbcl_scr_syn_totprob_r ~ allface_0b_rate + sex + interview_age + (1|site_id_l), data = data_0back[data_0back$network == "v1",])
lme4_0back_v1.M <- 
  lmer(dis ~ allface_0b_rate + sex + interview_age + (1|site_id_l), data = data_0back[data_0back$network == "v1",])
lme4_0back_v1.Y <- 
  lmer(cbcl_scr_syn_totprob_r ~ allface_0b_rate + dis + sex + interview_age + (1|site_id_l), data = data_0back[data_0back$network == "v1",])
lme4_med_0back_v1 <- 
  mediate(model.m = lme4_0back_v1.M, model.y = lme4_0back_v1.Y, treat='allface_0b_rate', mediator = 'dis', data=data_0back[data_0back$network == "v1",], sims=1000, dropobs=TRUE)

summary(lme4_0back_v1.0)
summary(lme4_0back_v1.M)
summary(lme4_0back_v1.Y)
summary(lme4_med_0back_v1) # ACME            -0.6463      -1.0603        -0.27  <2e-16 ***

## 2-back ###
# vis
lme4_2back_vis.0 <- 
  lmer(cbcl_scr_syn_totprob_r ~ allface_2b_rate + sex + interview_age + (1|site_id_l), data = data_2back[data_2back$network == "vis",])
lme4_2back_vis.M <- 
  lmer(dis ~ allface_2b_rate + sex + interview_age + (1|site_id_l), data = data_2back[data_2back$network == "vis",])
lme4_2back_vis.Y <- 
  lmer(cbcl_scr_syn_totprob_r ~ allface_2b_rate + dis + sex + interview_age + (1|site_id_l), data = data_2back[data_2back$network == "vis",])
lme4_med_2back_vis <- 
  mediate(model.m = lme4_2back_vis.M, model.y = lme4_2back_vis.Y, treat='allface_2b_rate', mediator = 'dis', data=data_2back[data_2back$network == "vis",], sims=1000, dropobs=TRUE)

summary(lme4_2back_vis.0)
summary(lme4_2back_vis.M)
summary(lme4_2back_vis.Y)
summary(lme4_med_2back_vis) # ACME            -1.1415      -1.6326        -0.66  <2e-16 ***

# lim
lme4_2back_lim.0 <- 
  lmer(cbcl_scr_syn_totprob_r ~ allface_2b_rate + sex + interview_age + (1|site_id_l), data = data_2back[data_2back$network == "lim",])
lme4_2back_lim.M <- 
  lmer(dis ~ allface_2b_rate + sex + interview_age + (1|site_id_l), data = data_2back[data_2back$network == "lim",])
lme4_2back_lim.Y <- 
  lmer(cbcl_scr_syn_totprob_r ~ allface_2b_rate + dis + sex + interview_age + (1|site_id_l), data = data_2back[data_2back$network == "lim",])
lme4_med_2back_lim <- 
  mediate(model.m = lme4_2back_lim.M, model.y = lme4_2back_lim.Y, treat='allface_2b_rate', mediator = 'dis', data=data_2back[data_2back$network == "lim",], sims=1000, dropobs=TRUE)

summary(lme4_2back_lim.0)
summary(lme4_2back_lim.M)
summary(lme4_2back_lim.Y)
summary(lme4_med_2back_lim) # ACME            -0.5729      -0.9724        -0.20  <2e-16 ***

# v1
lme4_2back_v1.0 <- 
  lmer(cbcl_scr_syn_totprob_r ~ allface_2b_rate + sex + interview_age + (1|site_id_l), data = data_2back[data_2back$network == "v1",])
lme4_2back_v1.M <- 
  lmer(dis ~ allface_2b_rate + sex + interview_age + (1|site_id_l), data = data_2back[data_2back$network == "v1",])
lme4_2back_v1.Y <- 
  lmer(cbcl_scr_syn_totprob_r ~ allface_2b_rate + dis + sex + interview_age + (1|site_id_l), data = data_2back[data_2back$network == "v1",])
lme4_med_2back_v1 <- 
  mediate(model.m = lme4_2back_v1.M, model.y = lme4_2back_v1.Y, treat='allface_2b_rate', mediator = 'dis', data=data_2back[data_2back$network == "v1",], sims=1000, dropobs=TRUE)

summary(lme4_2back_v1.0)
summary(lme4_2back_v1.M)
summary(lme4_2back_v1.Y)
summary(lme4_med_2back_v1) # ACME            -0.47589     -0.77228        -0.19   0.002 ** 
```
